\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-27 tex/paper.tex
save

    \item $\smash{\hat\theta}_f$ maximimizes the expected entropy
    $\mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$, which happens when all outcomes of
    $Z|f(X;\smash{\hat\theta}_f)$ are equally likely. In other words,
    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ since we
    require a uniform prior by construction. Note that in the continuous case,
    the supremum of the differential entropy over continuous distributions on
    the same bounded support is also realized by the uniform distribution over
    that support.
\end{itemize}
When the lower bound is active, we have $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
because of the second condition, which induces that the
optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal
quantity as a consequence of Proposition~\ref{prop:1}.

@@

\begin{figure*}
    \label{fig:toy}
    \begin{center}
        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
        \vspace{-0.8cm}
        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-1cm}
    \caption{Toy example.
    (Top left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
    (Top right) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ closer to $5$,
       hence explaining the dependency.
    (Bottom left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now closer to each other, indicating only a
       small dependency on $Z$.
    (Bottom right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function to erase the dependency on $Z$.
    }
\end{figure*}

\begin{figure}
    \label{fig:toy-training}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$ and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.}
\end{figure}

@@

Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build
a classifier $f$ minimizing $L_f(\theta_f)$ without adversarial training.
\glnote{to be continued}

>>>>

    \item $\smash{\hat\theta}_f$ maximizes the expected entropy
    $\mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$, which happens when all outcomes of
    $Z|f(X;\smash{\hat\theta}_f)$ are equally likely. In other words,
    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ since we
    require a uniform prior by construction. Note that in the continuous case,
    the supremum of the differential entropy over continuous distributions on
    the same bounded support is also realized by the uniform distribution over
    that support.
\end{itemize}
When the lower bound is active, we have $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
because of the second condition, which induces that the
optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal
quantity as a consequence of Proposition~\ref{prop:1}.

@@

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
        \vspace{-0.8cm}
        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-1cm}
    \caption{Toy example.
    (Top left) Conditional probability densities of the decision scores for $Z=0$, $2.5$ and $5$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
    (Top right) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ closer to $5$,
       hence explaining the dependency.
    (Bottom left) Conditional probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now closer to each other, indicating only a
       small dependency on $Z$.
    (Bottom right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function to erase the dependency on $Z$.
    }
    \label{fig:toy}
\end{figure*}

\begin{figure}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$
             and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.
             Adversarial training was performed for $1000$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
    \label{fig:toy-training}
\end{figure}

@@

Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
neural network classifier $f$ minimizing $L_f(\theta_f)$ without adversarial
training. As shown in the first row of Figure~\ref{fig:toy}, the resulting
classifier is not pivotal, as the conditional probability densities of its
decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
the nuisance. While not shown here, a classifier trained only from data
generated at the nominal value $Z=0$ would also not be pivotal.

By contrast, when using Algorithm~\ref{alg:adversarial-training} for training
$f$ jointly with a mixture density network $r$ modeling the conditional
distribution of the nuisance as observed through $f$, the resulting classifier
becomes effectively less dependent on $Z$, as shown in the second row of
Figure~\ref{fig:toy}. In particular, the conditional probability densities of
the decision scores $f(X;\theta_f)$ are now closer to each other, indicating only a
small  dependency on the nuisance, as theoretically expected.
\glnote{Discuss Figure~\ref{fig:toy-training}}
